What to simulate? Inferring the right direction for mental rotation

### 2014-04-29 man/cogsci-2014/cogsci-2014.tex
Updates in response to reviewer comments

\addtolength{\abovecaptionskip}{-0.25cm}
\addtolength{\belowcaptionskip}{-0.3cm}

@@

\begin{abstract}
  When people use mental imagery, how do they decide \textit{which}
  images to generate? To answer this question, we explored how mental
  simulation should be used in the classic psychological task of
  determining if two images depict the same object in different
  orientations \cite{Shepard1971}. Through a rational analysis of
  mental rotation, we formalized four models and compared them to
  human performance. We found that three models based on previous
  hypotheses in the literature were unable to account for several
  aspects of human behavior. The fourth, which is based on the idea of
  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, provides
  plausible explanations where the other models do not. Based on these
  results, we suggest that the question of ``what to simulate?'' is
  more difficult than has previously been assumed, and that an active
  learning approach holds promise for uncovering the answer.

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
    \caption{\textbf{Example stimuli and model behavior.} On the left
      is a ``flipped'' stimulus pair with a rotation of
      $120^\circ$. The grey plots show the actions that each model
      took for this stimulus. The color of each point (action)
      indicates whether the model's current mental image was generated
      under the hypothesis that the objects are the same (red) or
      flipped (blue). The similarity functions for each of these
      hypotheses are shown on the right.}
    \label{fig:model-traces}
  \end{center}
\end{figure*}

@@

The only models (of which the authors are aware) that seriously
attempted to address the decision of \textit{what} to simulate are
those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
models, the axis and direction of rotation are computed prior to
performing the rotation. One object is then rotated through the target
rotation, and is checked against the other object for
congruency. However, this approach assumes that the corresponding
points on the two objects can be easily identified, which is not
necessarily the case.  Indeed, the state-of-the-art in computer vision
suggests that there is more to this problem than simply checking for
congruency, particularly when shapes are complex or when the shapes
might not be exactly the same
\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally, recent
research shows that when performing \textit{physical} rotations,
people do not fully rotate until congruency is reached; they may even
rotate \textit{away} from initially near perfect matches
\cite{Gardony:2013gn}.

@@

Active sampling is the idea people gather new information in a manner
that increases certainty about the problem space. This notion has
recently gained support in several areas of cognitive science
\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
\cite{Juni:2011vo}. In the case of mental rotation, if people do not
know the minimum angle beforehand, then actively performing rotations
may be the best way to gather evidence about the similarity between
the observed shapes.

@@

Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
depict the same object, and $\hf$ is the hypothesis that the images
depict mirror-image objects. The posterior probability of each
hypothesis given the observed shapes is then: $p(h\ \vert\ X_a, X_b)
\propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta$,
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$. Because we ultimately want to determine
which hypothesis is more likely, the quantity of interest is a ratio
$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
(assuming that all rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(\hi)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(\hf)}.
  \label{eq:lh-ratio}
\end{equation}
If $\ell > 1$, then we accept the hypothesis that the images depict
the same object ($\hi$); if $\ell < 1$, then we accept the hypothesis
that the images depict flipped objects ($\hf$).

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so the similarity $S$ must
marginalize over the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Figure~\ref{fig:model-traces}.

Additionally, we assume that the observed shapes can only be
transformed by a small amount at a time, and each transformation takes
a non-negligible amount of time. Specifically, if the current mental
image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size.

@@

In order to approximate Equation~\ref{eq:lh-ratio} using samples of
the similarity function, we must decide \textit{which} places to
sample and \textit{when} stop sampling. The models below differ in how
they make these decisions; see Figure~\ref{fig:model-traces} for an
example.

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This gives us the true value of $\theta$, so
Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
test, where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
  \label{eq:mle-lh-ratio}
\end{equation}

@@

A model which does not know point correspondences could use the
following algorithm: (1) pick a random direction; (2) take a single
step; (3) if that step decreased similarity, then begin rotating in
the reverse direction, otherwise continue rotating in the original
direction; (4) continue rotating in the chosen direction until a
``match'' is found (defined as finding a value of $S$ that exceeds a
threshold); and (5) if no match was found, flip, and start over from
step one.  We assume that the locations where $S$ is greater than the
threshold correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, we use
Equation~\ref{eq:mle-lh-ratio}.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear RT found by \citeA{Shepard1971}. A second alternative is to
perform a Hill Climbing (\Hc{}) search; i.e., rotate in the direction
that increases similarity until no further improvement can be
found. In contrast with the \Th{} model, this results in arriving in a
\textit{local} maximum (which may or may not be the global
maximum). Thus, as with the \Oc{} and \Th{} models, we use
Equation~\ref{eq:mle-lh-ratio}.

@@

Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\ p(\hi)}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\ p(\hf)}.
\end{equation}
This distribution cannot easily be calculated; however, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
computing $p(\ell)$, we use $Z_D=Z_0-Z_1$ and compute:
$p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.  We then sample new observations
until we are at least 95\% confident that $Z_D\neq 0$. In other words,
when $p(Z_D<0)<0.025$, we accept $\hi$, and when $p(Z_D<0)>0.975$, we
accept $\hf$. To determine which points to sample, we can compute the
expected variance of $Z_h$ if we sampled a new observation at
$\theta_a$. From \citeA{Osborne:2012tm}, we compute the expected
variance as $\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
qc\theta_a))\dif\log S_h(\theta_a)$ for each of the actions in
Equation~\ref{eq:actions}, and then choose the action with the lowest
value.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Figure~\ref{fig:model-traces}). For each shape, we computed 20
``same'' and 20 ``flipped'' stimuli pairs, with 18 rotations
($\theta$) spaced at $20^\circ$ increments between $0^\circ$ and
$360^\circ$ (with $0^\circ$ and $180^\circ$ repeated twice, in order
to gather an equal number of responses for each angle between
$0^\circ$ and $180^\circ$). ``Same'' pairs were created by rotating
$\Xa$ by $\theta$; ``flipped'' pairs were first reflected $\Xa$ across
the $y$-axis, then rotated by $\theta$.

@@

Of the 247 participants, 200 (81\%) were included in our analyses. Of
the other 47, we excluded 10 (4\%) because of an experimental error, 6
(2.4\%) because they had already completed a related experiment, and
31 (12.6\%) because they failed a comprehension check, which was
defined as correctly answering at least 85\% of stimuli with a minimum
rotation of $0^\circ$ or $20^\circ$.  Additionally, we excluded 82
trials for which the RT was either less than 100ms or greater than
20s.

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
\Hc{} models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). This only had a major effect on the stopping criteria for the
\Bq{} model.

@@

As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This has
the interesting effect of actually causing the \Th{} model to
\textit{over}rotate (see Figure~\ref{fig:model-traces}). This is
because the step size is large enough that it sometimes misses the
global maximum, and must do another full rotation to find it. However,
when the step size is smaller, the fit to human data is slightly
worse.

@@

Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ had a fairly significant effect
(Figure~\ref{fig:response-time-accuracy}). With just a small bias of
$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
``flipped'' stimuli; this separation is similar to the trend also
observed in human RTs. While the prior also had an effect on accuracy,
this effect did not reflect the trends in human behavior.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
    \caption{\textbf{Typical RT curves for a single shape.}  These
      plots correspond to the shape shown in
      Figure~\ref{fig:model-traces}. Left: human curves are either
      linear (as with the ``same'' pairs), or linear and then flat (as
      with the ``flipped'' pairs). Middle: the \Th{} model does not
      have a reliably monotonic relationship with rotation. Right: the
      \Bq{} model is roughly linear.}
    \label{fig:response-time-stimulus}
  \end{center}
\end{figure}

@@

% \section{Acknowledgments}
% {\small This research was supported by ONR MURI grant number
% N00014-13-1-0341, and a Berkeley Fellowship awarded to JBH. }

>>>>

\addtolength{\abovecaptionskip}{-0.25cm}
\addtolength{\belowcaptionskip}{-0.35cm}

@@

\begin{abstract}
  When people use mental imagery, how do they decide \textit{which}
  images to generate? To answer this question, we explored how mental
  simulation should be used in the classic psychological task of
  determining if two images depict the same object in different
  orientations \cite{Shepard1971}. Through a rational analysis of
  mental rotation, we formalized four models and compared them to
  human performance. We found that three models based on previous
  hypotheses in the literature were unable to account for several
  aspects of human behavior. The fourth is based on the idea
  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, which is a
  strategy of choosing actions that will provide the most
  informative. This last model provides a plausible account of how
  people use mental rotation, where the other models do not. Based on
  these results, we suggest that the question of ``what to simulate?''
  is more difficult than has previously been assumed, and that an
  active learning approach holds promise for uncovering the answer.

@@

@@

The only models (of which the authors are aware) that seriously
attempted to address the decision of \textit{what} to simulate are
those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
models, the axis and direction of rotation are computed prior to
performing the rotation. One object is then rotated through the target
rotation, and is checked against the other object for
congruency. However, this approach assumes that the corresponding
points on the two objects can be easily identified, which is not
necessarily the case.  Indeed, the state-of-the-art in computer vision
suggests that there is more to this problem than checking for
congruency, particularly when the shapes are complex or not exactly
the same \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally,
recent research shows that when performing \textit{physical}
rotations, people do not rotate until congruency is reached; they may
even rotate \textit{away} from initially near perfect matches
\cite{Gardony:2013gn}.

@@

Active sampling is the idea people gather new information in a manner
that increases certainty about the problem space. An everyday example
of this can be observed in the game of ``20 questions'', in which one
person thinks of a concept, and another has to guess the concept in 20
questions or less. The first question is almost always ``person,
place, or thing?'', because the answer provides the most possible
information about the concept of interest. Active sampling has
recently gained support across several areas of cognitive science
\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
\cite{Juni:2011vo}. In the case of mental rotation, actively choosing
rotations may be the best way to gather evidence about the similarity
between the observed shapes when the angle of rotation is unknown.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/stimuli_and_similarity.pdf}
    \caption{\textbf{Example stimuli and similarities.} This figure
      shows a ``flipped'' stimulus pair with a rotation of
      $120^\circ$, and the corresponding similarity functions for each
      hypothesis. Arrows indicate where each shape lies on the curve.}
    \label{fig:shapes}
  \end{center}
\end{figure}

The task we are interested in modeling involves observing two images
and determining whether one image depicts the ``same'' object as the
other image (differing only by a rotation), or a ``flipped'' version
of the object in the other image (differing by a reflection and then a
rotation).

Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
depict the same object, and $\hf$ is the hypothesis that the images
depict mirror-image objects. The posterior probability of each
hypothesis given the observed shapes is then: $p(h\ \vert\ X_a, X_b)
\propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta$,
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$. Because we ultimately want to determine
which hypothesis is more likely, the quantity of interest is a ratio
$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
(assuming that all rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
  \label{eq:lh-ratio}
\end{equation}
where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
we accept the hypothesis that the images depict the same object
($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
depict flipped objects ($\hf$).

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so the similarity $S$ must
marginalize over the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Figure~\ref{fig:shapes}.

Additionally, we assume that the observed shapes can only be
transformed by a small amount at a time, and each transformation takes
a non-negligible amount of time. Specifically, if the current mental
image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset to $0^\circ$, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size.

@@

In order to approximate Equation~\ref{eq:lh-ratio} using samples of
the similarity function, we must decide \textit{which} places to
sample and \textit{when} stop sampling. The models below differ in how
they make these decisions.

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This gives us the true value of $\theta$, so
Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
test, where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)\cdot{}p_0}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)\cdot{}p_1}.
  \label{eq:mle-lh-ratio}
\end{equation}

If we give equal weight to the two hypotheses, then the priors cancel
out; if we weigh one hypothesis more heavily, then our decision will
be biased towards that hypothesis. However, unless the likelihood
ratio is already very close to 1, small biases in the prior will not
make much of a difference as to whether we accept $\hi$ or $\hf$.

@@

A model which does not know point correspondences could use the
following algorithm: (1) pick a random direction; (2) take a single
step; (3) if that step decreased similarity, then begin rotating in
the reverse direction, otherwise continue rotating in the original
direction; (4) continue rotating in the chosen direction until a
``match'' is found (defined as finding a value of $S$ that exceeds a
threshold); and (5) if no match was found, flip, and start over from
step one.  We assume that the locations where $S$ is greater than the
threshold correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, we use
Equation~\ref{eq:mle-lh-ratio}. Additionally, we only allow for the
``flip'' action after no match has been found, because there is no
particularly principled way for the \Th{} model to choose when to
flip.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear RT found by \citeA{Shepard1971}. A second alternative is to
perform a Hill Climbing (\Hc{}) search; i.e., rotate in the direction
that increases similarity until no further improvement can be
found. In contrast with the \Th{} model, this results in arriving in a
\textit{local} maximum (which may or may not be the global
maximum). Thus, as with the \Oc{} and \Th{} models, we use
Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip'' action
after a local maximum has been reached, because like the \Th{} model,
there is otherwise no principled way for the \Hc{} model to choose
when to flip.

@@

Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\cdot{}p_0}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\cdot{}p_1}.
\end{equation}

This distribution cannot easily be calculated; however, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Instead, we use
$Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
$p(Z_D)\propto\mathcal{N}(p_0\cdot{}\mathbb{E}[Z_0] -
p_1\cdot{}\mathbb{E}[Z_1], p_0^2\cdot{}\mathbb{V}(Z_0) +
p_1^2\cdot{}\mathbb{V}(Z_1))$.
We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. Because we compare
the hypotheses in order to determine when to stop sampling, biasing
the prior towards one hypothesis should result in requiring less
evidence for that hypothesis before stopping, and more evidence for
the other hypothesis. This is in contrast to the previous models,
where we would not expect the prior to have much effect.

To determine which points to sample, we can compute the expected
variance of $Z_h$ if we sampled a new observation at $\theta_a$. From
\citeA{Osborne:2012tm}, we compute the expected variance as
$\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
qc\theta_a))\dif\log S_h(\theta_a)$
for each of the actions in Equation~\ref{eq:actions}, and then choose
the action with the lowest value.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Figure~\ref{fig:shapes}). For each shape, we computed 20 ``same'' and
20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced at
$20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
number of responses for each angle between $0^\circ$ and
$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
$y$-axis, then rotated by $\theta$.

@@

Of the 247 participants, 200 (81\%) were included in our analyses. Of
the other 47, we excluded 10 (4\%) because of an experimental error, 6
(2.4\%) because they had already completed a related experiment, and
31 (12.6\%) because they failed a comprehension check, which was
defined as correctly answering at least 85\% of stimuli with a
rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.
Additionally, we excluded 82 trials for which the RT was either less
than 100ms or greater than 20s.

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
\Hc{} models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). As expected, this only had a major effect on the stopping
criteria for the \Bq{} model.

@@

As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This has
the interesting effect of actually causing the \Th{} model to
\textit{over}rotate. This is because the step size is large enough
that it sometimes misses the global maximum, and must do another full
rotation to find it. However, when the step size is smaller, the fit
to human data is slightly worse.

@@

Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ had a fairly significant effect
(Figure~\ref{fig:response-time-accuracy}). As expected, with just a
small bias of $p(h=0)=0.55$, there was a clear separation in RTs for
``same'' versus ``flipped'' stimuli: because of this bias, the model
needed less evidence before accepting $h=0$ (thus taking less
time). This separation is similar to the trend also observed in human
RTs. The prior also had an effect on accuracy: the bias towards $h=0$
meant that the model was more likely to judge a pair as ``same'',
thus, accuracy increased for ``same'' pairs, but decreased for
``flipped'' pairs. However, this effect on accuracy did not reflect
the trends in human behavior.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
    \caption{\textbf{Typical RT curves for a single object.}  These
      plots correspond to the object shown in
      Figure~\ref{fig:shapes}. Left: human curves are either linear
      (as with the ``same'' pairs), or linear and then flat (as with
      the ``flipped'' pairs). Middle: the \Th{} model does not have a
      reliably monotonic relationship with rotation. Right: the \Bq{}
      model is roughly linear.}
    \label{fig:response-time-stimulus}
  \end{center}
\end{figure}

@@

\section{Acknowledgments} {\small This research was supported by ONR
  MURI grant number N00014-13-1-0341, and a Berkeley Fellowship
  awarded to JBH. }
